### 2015-12-11 paper/1-introduction.tex
Make space by removing C++11 background

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
And while \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, this is often insufficient.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to \textsf{DEVS} simulation.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The exact reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other in specific models.
The applicability of parallel simulation capabilities of current tools, is therefore limited to the domain of simulation models.

Users that wish to simulate a wide variety of models, with different ideal synchronization protocols, are therefore out of luck:
either they accept the lower performance for some of the models, or they use two different simulation kernels.
Neither is acceptable for the simulation of complex models: performance can decrease to unacceptable levels, and simulation kernel APIs can diverge significantly.

In this paper, we introduce DEVS-Ex-Machina (``dxex''), our simulation tool which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparant to the simulated model.
Users should merely determine, at the start of simulation, which protocol they wish to use.
Our tool is implemented in C++11, to increase both performance and portability across different platforms.

@@

Dxex, as well as all of our benchmark models, results, and profiling results, can be found online at \url{https://bitbucket.org/bcardoen/devs-ex-machina}.

The remainder of this paper is organized as follows:
Section~\ref{sec:2-background} introduces the necessary background on synchronization protocols and our implementation language.
Section~\ref{sec:3-features} elaborates on some of our features, and our design that enables our flexibility.
In Section~\ref{sec:4-performance}, we evaluate performance of our tool by comparing its different synchronization protocols, as well as a comparison to adevs.
Related work is discussed in Section~\ref{sec:5-related-work}.
Section~\ref{sec:6-conclusion} concludes the paper and gives future work.

>>>>

But with the ever increasing complexity of simulation models, parallel simulation becomes necessary to perform the simulation within reasonable time bounds.
And while \textsf{Parallel DEVS}~\cite{ParallelDEVS} was introduced to increase parallelism, this is often insufficient.
Several synchronization protocols from the discrete event simulation community~\cite{FujimotoBook} have been applied to \textsf{DEVS} simulation.
While several parallel \textsf{DEVS} simulation kernels exist, they are often limited to a single synchronization protocol.
The reason for different synchronization protocols, however, is that their distinct nature makes them applicable in different situations, each outperforming the other in specific models.
The applicability of parallel simulation capabilities of current tools is therefore limited.

Users who simulate a wide variety of models, with different ideal synchronization protocols, are therefore out of luck:
either they accept lower performance for some models, or they use different simulation kernels.
Neither is acceptable for complex models: performance can decrease to unacceptable levels, and simulation kernel APIs can diverge significantly.

This paper introduces DEVS-Ex-Machina (``dxex''), our simulation tool which offers multiple synchronization protocols: no synchronization (sequential execution), conservative synchronization, or optimistic synchronization.
The selected synchronization protocol is transparant to the simulated model.
Users should merely determine, at the start of simulation, which protocol they wish to use.
Our tool is based on PythonPDEVS, but implemented in C++11, to increase both performance and portability across different platforms.

@@

Dxex can be found online at \url{https://bitbucket.org/bcardoen/devs-ex-machina}.

The remainder of this paper is organized as follows:
Section~\ref{sec:2-background} introduces the necessary background on synchronization protocols.
Section~\ref{sec:3-features} elaborates on some of our features, and our design that enables our flexibility.
In Section~\ref{sec:4-performance}, we evaluate performance of our tool by comparing its different synchronization protocols, as well as a comparison to adevs.
Related work is discussed in Section~\ref{sec:5-related-work}.
Section~\ref{sec:6-conclusion} concludes the paper and gives future work.

### 2015-12-11 paper/2-background.tex
Make space by removing C++11 background

This section briefly introduces two distinct synchronization protocols, as used by dxex.
Furthermore, we make note of several new features of C++11.

@@

\subsection{C++11 Parallelism Features}
Apart from various other additions, C++11 adds a wide range of portable synchronization primitives to the Standard Library.
In earlier versions of the language standard, one had to resort to non-portable C implementations.
Furthermore, C++11 is the first version of the standard that explicitly defines a multi-threaded abstract machine memory model.
Most modern compilers support the full standard, allowing the kernels to be portable by default on any standard compliant platform.

Simulation values that need to be shared between nodes, can now be coded portable and efficiently through the use of atomic fields.
This avoids latencies caused by the exchange of synchronization messages.
For an in-depth study, we refer to~\cite{CPE:CPE3007}.

>>>>

This section briefly introduces the two synchronization protocols used by dxex: conservative and optimistic synchronization.

@@
